Modeling the Circumnuclear Mediums and Accetion Rates of Quiescent Supermassive Black Holes

### 2014-12-16 ms.tex
2.2 second pass

The source term in the continuity equation (eq.~[\ref{eq:drhodt}]),
\begin{align}
  q=\frac{\eta \rhostar}{\th},
\label{eq:q}
\end{align}
represents mass input from stellar winds, which we parameterize in
terms of the fraction $\eta$ of the stellar density $\rhostar$ being
recycled into gas on the Hubble time $\th = 1.4 \times 10^{10}$
years.  To good approximation $\eta\simeq 0.02 (t_{\star}/t_h)^{-1.3}$ at time $t_{\star}$ following a burst of star formation (e.g., \citet{Ciotti+91}),\footnote{Note that 
\citet{Ciotti+91} gives the overall mass return rate from evolved
stars as a function of B-band luminosity. We give the mass return rate
per unit volume. However, our expressions are equivalent.} although $\eta$ will generally be higher in cases of ongoing star formation.

Source terms $\propto q$ also appear in the momentum and entropy equations (eqs.~[\ref{eq:dvdt}] and [\ref{eq:dsdt}]) due to the fact that the isotropic mass injection in the rest frame of the SMBH
represents a source of momentum and energy relative to the mean flow (see also below).  Physically these terms result from the mismatch between the properties of virialized gas injected locally by stellar winds and the mean properties of the background flow.   As we shall discuss in $\S$, the cooling term $\propto q p/\rho$ represents an important stabilizing term in comparison to radiative cooling.

The last term in equation (\ref{eq:dsdt}) accounts for heat transportation by electron conductivity, where $\kappa_{\rm eff}=\kappa_{\rm spitz}/(1+\psi)$ (\citealt{DaltonBalbus:1993a}) is the conducitivity and $\kappa_{\rm spitz} = \kappa_0 T^{5/2}$ is the classical \citet{Spitzer62} value ($\kappa_0\simeq 2\times 10^{-6}$ in cgs units).  The flux limitor $\psi = \kappa_{\rm spitz} \nabla T/(5\phi \rho c_s^3)$ saturates the conductive flux if the mean free path for electron coulomb scattering becomes comparable to the temperature length scale.  Here $c_s \equiv (kT/\mu m_p)^{1/2}$ is the isothermal sound speed and $\phi$ is an uncertain dimensionless constant, typically taken to be of order unity of less (we adopt $\phi = XXX$).  Even a weak magnetic field could in principle also suppress the conductivity perpendicular to the field.  However, the radially-decreasing temperature profiles of interest are susceptible to the magneto-thermal instability (\citealt{Balbus01}), the non-linear evolution of which is believed to result in a radially-directed field structure (\citealt{Parrish&Stone07}).

Following \citet{ShcherbakovWong+:2014a}, the term $\propto \vw^2 = \sigma(r)^2+v_{w,0}^2$ in the entropy equation accounts heating sources.  Here $\sigma(r) = \sqrt{GM_{\rm enc}/r}$ is the velocity dispersion of the stars, accounting for the minimum amount of shock heating from the stellar winds due to the random movement of stars with respect to each other at the characteristic velocity $\sim \sigma$.  This heating is present even for low wind velocities $\ll \sigma$, as characterize the majority of AGB mass loss.
The term $v_{w,0}^{2}$ parameterizes the contribution of additional sources of heating, including main sequence winds, millisecond pulsars, supernovae, AGN feedback, etc.  As discussed in $\S\ref{sec:heating}$, the value of $v_{w,0}$ is uncertain and will in general depend on the SMBH mass and the age of the stellar population.

Equations (\ref{eq:drhodt})-(\ref{eq:dsdt}) are solved using a sixth
order finite difference scheme with a third order Runge-Kutta scheme
for time integration and artificial viscosity terms in the velocity
and entropy equations for numerical stability (\citealt{Brandenburg:2003a}).  We solve the equations for
different choices of $v_{\rm w,0} = 300, 600, 1200$ km s$^{-1}$ spanning the physically expected
range.  Although we are most interested in the steady-state CNM structure, the time-dependent equations are solved to avoid numerical issues that arise near the critical sonic points, of which there are up to two because of the supersonic inflow on small radial scales and a possible supersonic outflow on large scales.

Radiative cooling is not included in the entropy equation, although its effect is assessed critically in $\S\ref{sec:cooling}$.  Across much of the parameter space radiative cooling is subdominant compared to other sources of cooling (e.g., advection).  Radiative cooling cannot be neglected in some cases (if
$\vwO$ is low and/or $\eta$ is high), once it becomes important it leads to thermal instabilities, which our one-dimensional time-dependent modeal cannot account for anyways (e.g.~\citealt{McCourt+12}).  A chief goal of this work is instead to use our steady state solutions neglecting radiation to ascertain over what physical conditions cooling instabilities are likely to develop ($\S\ref{sec:cooling}$).

>>>>

The source term in the continuity equation (eq.~[\ref{eq:drhodt}]),
\begin{align}
  q=\frac{\eta \rhostar}{\th},
\label{eq:q}
\end{align}
represents mass input from stellar winds, which we parameterize in
terms of the fraction $\eta$ of the stellar density $\rhostar$ being
recycled into gas on the Hubble time $\th = 1.4 \times 10^{10}$
years.  To good approximation $\eta\simeq 0.02 (t_{\star}/t_h)^{-1.3}$ at time $t_{\star}$ following a burst of star formation (e.g., \citealt{Ciotti+91}),\footnote{\citet{Ciotti+91} give the overall mass return rate from evolved stars as a function of B-band luminosity.  Although we give the mass return rate per unit volume, our expressions are equivalent.} although $\eta$ is higher in cases of ongoing star formation.

Source terms $\propto q$ also appear in the momentum and entropy equations (eqs.~[\ref{eq:dvdt}] and [\ref{eq:dsdt}]) because isotropic mass injection in the SMBH rest frame represents a source of momentum and energy relative to the mean flow.  Physically, these result from the mismatch between the properties of virialized gas injected locally by stellar winds and the mean properties of the background flow.   The term $\propto -q c_{s}^{2}$ turns out to be an important in stabilizing the flow to runaway cooling ($\S$).

The last term in equation (\ref{eq:dsdt}) accounts for heat transportation by electron conductivity, where $\kappa_{\rm eff}=\kappa_{\rm spitz}/(1+\psi)$ (\citealt{DaltonBalbus:1993a}) is the conducitivity and $\kappa_{\rm spitz} = \kappa_0 T^{5/2}$ is the classical \citet{Spitzer62} value ($\kappa_0\simeq 2\times 10^{-6}$ in cgs units).  The flux limitor $\psi = \kappa_{\rm spitz} \nabla T/(5\phi \rho c_s^3)$ saturates the conductive flux if the mean free path for electron coulomb scattering becomes comparable to the temperature length scale, where $c_s \equiv (kT/\mu m_p)^{1/2}$ is the isothermal sound speed and $\phi \lesssim 1$ is an uncertain dimensionless constant (we adopt $\phi = XXX$).  Even a weak magnetic field perpendicular to the flow could in principle suppress the conductivity.  However, the radially-decreasing temperature profiles of interest are susceptible to the magneto-thermal instability (\citealt{Balbus01}), the non-linear evolution of which is believed to result in a radially-directed field structure (\citealt{Parrish&Stone07}).

The term $\propto \vw^2 = \sigma(r)^2+v_{w,0}^2$ in the entropy equation accounts for external heating sources (e.g., \citealt{ShcherbakovWong+:2014a}).  The first term accounts for the minimum shock heating of the stellar winds due to the random movement of stars with respect to each other at the characteristic velocity $\sim \sigma = \sqrt{GM_{\rm enc}/r}$, as is present even for the low wind velocities $\ll \sigma$ that characterize AGB mass loss in old stellar populations.  The second term, $v_{w,0}^{2}$, parameterizes the contribution of additional sources of heating, such as faster main sequence or Wolf-Rayet winds, millisecond pulsars, supernovae, AGN feedback, etc.  As discussed in $\S\ref{sec:heating}$, the value of $v_{w,0}$ is uncertain and will in general depend on the SMBH mass and the age of the stellar population.

Equations (\ref{eq:drhodt})-(\ref{eq:dsdt}) are solved using a sixth
order finite difference scheme with a third order Runge-Kutta scheme
for time integration and artificial viscosity terms in the velocity
and entropy equations for numerical stability (\citealt{Brandenburg:2003a}).  We solve the equations for different choices of $v_{\rm w,0} = 300, 600, 1200$ km s$^{-1}$ spanning a physically plausible
range.  Although we are interested in the steady-state intflow/outflow structure, the time-dependent equations are solved to avoid numerical issues that arise near the critical sonic points, of which there are up to two because of the supersonic inflow on small radial scales and a possible supersonic outflow on large scales.

Radiative cooling is not included in our calculation.  Across much of the parameter space radiative cooling is subdominant compared to other sources of cooling.  Radiative cooling cannot be neglected when $\vwO$ is low (and/or $\eta$ is high), but once important it leads to thermal instabilities (e.g.~\citealt{McCourt+12}) which our 1D time-independent model cannot account for self-consistently anyways.  Indeed, one of the main goals of this work is to use our analytic and numerical solutions neglecting radiation to ascertain over what physical conditions cooling instabilities are likely to develop ($\S\ref{sec:cooling}$).
